---
title: "Maloney_HW1"
author: "Emily Maloney"
date: "January 22, 2019"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Homework - Chapters 2 and 3

```{r}
library(tidyverse)
library(rethinking)
library(tidybayes)
```

##Chapter 2 Homework

###Medium Problems
####2M1
```{r}
#WWW
              #define grid
              p_grid <- seq(from = 0, to = 1, length.out = 100)
              
              #define prior
              prior <- rep(1, 100)
              #prior <- exp(-5*abs(p_grid - 0.5))
              
              #compute likelihood at each value in the grid
              lh <- dbinom(3, size = 3, prob = p_grid)
              
              #compute product of lh & prior
              upost <- lh * prior
              
              #standardize post
              post <- upost/sum(upost)
              
              plot(p_grid, post, type = "b")

    #WWWL
              #define grid
              p_grid <- seq(from = 0, to = 1, length.out = 100)
              
              #define prior
              prior <- rep(1, 100)
              #prior <- exp(-5*abs(p_grid - 0.5))
              
              #compute likelihood at each value in the grid
              lh <- dbinom(3, size = 4, prob = p_grid)
              
              #compute product of lh & prior
              upost <- lh * prior
              
              #standardize post
              post <- upost/sum(upost)
              
              plot(p_grid, post, type = "b")
              
    #LWWLWWW
              #define grid
              p_grid <- seq(from = 0, to = 1, length.out = 100)
              
              #define prior
              prior <- rep(1, 100)
              #prior <- exp(-5*abs(p_grid - 0.5))
              
              #compute likelihood at each value in the grid
              lh <- dbinom(5, size = 7, prob = p_grid)
              
              #compute product of lh & prior
              upost <- lh * prior
              
              #standardize post
              post <- upost/sum(upost)
              
              plot(p_grid, post, type = "b")
```

####2M2
```{r}
#WWW
              #define grid
              p_grid <- seq(from = 0, to = 1, length.out = 100)
              
              #define prior
              prior <- ifelse(p_grid < 0.5, 0, 2)
              
              #compute likelihood at each value in the grid
              lh <- dbinom(3, size = 3, prob = p_grid)
              
              #compute product of lh & prior
              upost <- lh * prior
              
              #standardize post
              post <- upost/sum(upost)
              
              plot(p_grid, post, type = "b")
              
              #WWWL
              #define grid
              p_grid <- seq(from = 0, to = 1, length.out = 100)
              
              #define prior
              prior <- ifelse(p_grid < 0.5, 0, 2)
              
              #compute likelihood at each value in the grid
              lh <- dbinom(3, size = 4, prob = p_grid)
              
              #compute product of lh & prior
              upost <- lh * prior
              
              #standardize post
              post <- upost/sum(upost)
              
              plot(p_grid, post, type = "b")
              
              #LWWLWWW
              #define grid
              p_grid <- seq(from = 0, to = 1, length.out = 100)
              
              #define prior
              prior <- ifelse(p_grid < 0.5, 0, 2)
              #prior <- exp(-5*abs(p_grid - 0.5))
              
              #compute likelihood at each value in the grid
              lh <- dbinom(5, size = 7, prob = p_grid)
              
              #compute product of lh & prior
              upost <- lh * prior
              
              #standardize post
              post <- upost/sum(upost)
              
              plot(p_grid, post, type = "b")
```

####2M3
```{r}
pwe <- 0.7
plm <- 1
lh <- 0.3/1
priorodds <- 1
postodds <- priorodds*lh
(post <- postodds/(postodds + 1))
```
The probability of the globe was Earth, given seeing land is `r post`.

####2M4
```{r}
ww <- 0 
bw <- 1 
bb <- 2

(p <- bb/(bb + bw + ww))
```
Given the fact that a card with two white sides cannot produce a black side facing up, a black and white card can produce a black side facing up 1 way, and a card with two black sides can produce a black side facing up two ways, the probablity that the other side of a card with a black face up is 2/3. 

####2M5
```{r}
ww <- 0 
bw <- 1 
bb <- 2*2

(p <- bb/(bb + bw + ww))
```
If there are two cards that have black on both sides, the probability that a card with a black side facing up also has black on the other side is now 4/5. 

####2M6
```{r}
ww <- 0*3
bw <- 1*2 
bb <- 2

(p <- bb/(bb + bw + ww))
```
If there are two ways to pull out a black and white card and 3 ways to pull out a white and white card for every way to pull out a black and black card, the probability that the other side is black is now 1/2. 

####2M7
```{r}
ww <- 0 * (1 + 0)
bb <- 2 * (2 + 1)
bw <- 1 * (0 + 2)

(p <- bb/(bb + bw + ww))
```
If a second card is drawn with a white side face up, the probability that the first card with the black side facing up has black on the other side, is now 0.75. 

## Chapter 3 Homework
###Easy Problems


```{r data}
p_grid <- seq(from = 0, to = 1, length.out = 1000)
prior <- rep(1, 1000)
lh <- dbinom(6, size = 9, prob = p_grid)
post <- lh * prior
post <- post/sum(post)
set.seed(100)
samples <- sample(p_grid, prob = post, size = 1e4, replace = T)
samples <- as_tibble(samples)
```

####3E1

```{r}
less20 <- samples %>% filter(value < 0.2) %>%
            summarise(sum = n()/1e4)
```

There is 0.005 posterior probability that lies below p = 0.2. 

####3E2
```{r}
more80 <- samples %>% filter(value > 0.8) %>%
            summarise(sum = n()/1e4)
```
There is 0.0112 posterior probability that lies above p = 0.8. 

####3E3
```{r}
bw2080 <- samples %>% filter(value > 0.2 & samples < 0.8) %>%
            summarise(sum = n()/1e4)
```
There is 0.0888 posterior probability that lies between p = 0.2 and p = 0.8. 

####3E4
```{r}
quantile(samples$value, p = 0.2)
```
20% of the posterior probability lies below p = 0.52. 

####3E5
```{r}
quantile(samples$value, p = 0.8)
```
20% of the posterior probability lies above p = 0.757. 

####3E6
```{r}
HPDI(samples$value, p = 0.66)
```
The values of p which contain the narrowest interval equal to 66% of the posterior probability are 0.521 and 0.785. 

####3E7
```{r}
quantile(samples$value, p = 0.83)
quantile(samples$value, p = 0.17)
```
The values of p which contain 66% of the posterior probability, assuming equal posterior probability both below and above the interval are 0.500 and 0.769. 

###Medium Problems

####3M1
```{r}
n <- 1000
n_success <- 8
n_trials <- 15

d <- tibble(p_grid = seq(from = 0, to = 1, length.out = n),
            prior = 1) %>%
     mutate(lh = dbinom(n_success, size = n_trials, prob = p_grid),
            post = lh * prior,
            post = post/sum(post))

d %>%
            ggplot(aes(x = p_grid, y = post)) +
            geom_line(size = 1/10) + 
            labs(x = "p(water)",
                 y = "posterior probability") + ggtitle("8 Water in 15 Tosses")
```

####3M2
```{r}
#drawing samples
samples <- tibble(samples = sample(d$p_grid, prob = d$post, size = 10000, replace = T)) %>%
           mutate(sample_n = 1:n())
head(samples)

#90% HPDI
HPDI(samples$samples, p = 0.9)
```
The 90% HPDI for p is 0.338-0.731. 

####3M3
```{r}
ppc <- tibble(sample = rbinom(1e4, size = 15, prob = samples$samples))
p8 <- ppc %>% filter(sample == 8) %>%
              summarise(sum = n()/1e4)

```
There is  a `p8` probability of getting 8 tosses out of 15. 

####3M4
```{r}
newsim <- tibble(sample = rbinom(1e4, size = 9, prob = samples$samples))
p69 <- newsim %>% filter(sample == 6) %>%
                  summarise(sum = n()/1e4)
```
There is a `p69` probability of getting 6 tosses out of 9. 

####3M5
```{r}
n <- 1000
n_success <- 8
n_trials <- 15

d <- tibble(p_grid = seq(from = 0, to = 1, length.out = n),
            prior = ifelse(p_grid < 0.5, 0, 2)) %>%
     mutate(lh = dbinom(n_success, size = n_trials, prob = p_grid),
            post = lh * prior,
            post = post/sum(post))

d %>%
            ggplot(aes(x = p_grid, y = post)) +
            geom_line(size = 1/10) + 
            labs(x = "p(water)",
                 y = "posterior probability") + ggtitle("8 Water in 15 Tosses, new Prior")

#drawing samples
samples <- tibble(samples = sample(d$p_grid, prob = d$post, size = 10000, replace = T)) %>%
           mutate(sample_n = 1:n())

#90% HPDI
HPDI(samples$samples, p = 0.9)

#posterior predictive check
ppc <- tibble(sample = rbinom(1e4, size = 15, prob = samples$samples))
p8 <- ppc %>% filter(sample == 8) %>%
              summarise(sum = n()/1e4)
print(p8)

#probability of observing 6 water in 9 tosses
newsim <- tibble(sample = rbinom(1e4, size = 9, prob = samples$samples))
p69 <- newsim %>% filter(sample == 6) %>%
                  summarise(sum = n()/1e4)
print(p69)
```

